Covariate analysis of UFL-Mayo-ISB RNASeq (both cerebellum and temporal cortex)

### 2016-05-16 code/Rmd/Mayo_RNASeq_CellMarkersAdjusted.Rmd
Minor changes

### Cell type markers (i.e., eigen genes of cell types) as additional covariates
```{r cellMark}
load(synGet('syn4893059')@filePath)
GeneSets = GeneSets[grep("Zhang", names(GeneSets))] %>%
  ldply(.fun = function(x){
    data.frame(hgnc_symbol = x)
  }, .id = 'CellType') %>%
  dplyr::mutate(CellType = gsub("Zhang:","",CellType)) %>%
  left_join(Ensemble2HGNC) %>%
  filter(!is.na(ensembl_gene_id))

@@

p = METADATA %>%
  dplyr::select(Diagnosis, Tissue, one_of(colnames(COVARIATES.cellmarkers))) %>%
  gather(CellType, EigenGeneExpression, -ID, -Diagnosis, -Tissue) %>%
  ggplot(aes(x = Diagnosis, y = abs(EigenGeneExpression), fill = Tissue)) + geom_boxplot() + facet_grid(.~CellType)
p                         
```

@@

pl = list()
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=factor(Source), color=as.factor(Tissue), size=as.numeric(OPC)))
p <- p + theme_bw() + theme(legend.position="top") +scale_shape_manual(values = 1:2)
# p <- p + geom_text(aes(label= ID), size=4, hjust=0)
pl[[1]] <- p

p <- ggplot(plotdata, aes(x=PC3, y=PC4))
p <- p + geom_point(aes(shape=factor(Source), color=as.factor(Diagnosis), size=as.numeric(OPC)))
p <- p + theme_bw() + theme(legend.position="top") +scale_shape_manual(values = 1:2)
# p <- p + geom_text(aes(label= ID), size=4, hjust=0)
pl[[2]] <- p

multiplot(plotlist = pl, cols = 1)
```
```{r decompse.normalise.data2, fig.height=9, fig.width=25, results='asis'}
# Tree based clustering
sampleTrees = flashClust::hclust(dist(t(VOOM.GENE_EXPRESSION$E)), method = "average")
tree = WGCNA::cutreeStatic(sampleTrees, cutHeight = 135, minSize=10)

@@

### Clustering residual gene expression
```{r decompse.normalise.data3, fig.height=10, fig.width=9, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION[,-(1)], scale.=T)

@@

# Get differnetial expression
DE = list()
p = list()
for (i in 1:12){
  DE[[i]] = topTable(FIT.CONTR, coef=i, number = dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1])
  DE[[i]] = rownameToFirstColumn(DE[[i]], 'ensembl_gene_id') %>%
    left_join(Ensemble2HGNC)
  p[[i]] <- ggplot(DE[[i]], aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
  p[[i]] <- p[[i]] + geom_hline(yintercept = -log10(0.05), color = 'red') 
  p[[i]] <- p[[i]] + ggtitle(gsub('BrainRegion.Diagnosis','',colnames(contrast)[i]))
}
multiplot(plotlist=p, cols = 3)
names(DE) = gsub('BrainRegion.Diagnosis','',colnames(contrast))

DE = plyr::ldply(DE, .id = 'Comparison')
tmp = DE %>% 
  dplyr::filter(adj.P.Val <= 0.05) %>%
  group_by(Comparison) %>%
  dplyr::summarize(nDiffExpGenes = length(unique(hgnc_symbol)))
kable(tmp)
```

>>>>

### Cell type markers (i.e., eigen genes of cell types) as additional covariates
```{r cellMark, fig.height=10, fig.width = 15}
load(synGet('syn4893059')@filePath)
GeneSets = GeneSets[grep("Zhang", names(GeneSets))] %>%
  ldply(.fun = function(x){
    data.frame(hgnc_symbol = x)
  }, .id = 'CellType') %>%
  dplyr::mutate(CellType = gsub("Zhang:","",CellType)) %>%
  left_join(Ensemble2HGNC) %>%
  filter(!is.na(ensembl_gene_id))

@@

p = METADATA %>%
  dplyr::select(Diagnosis, Tissue, one_of(colnames(COVARIATES.cellmarkers))) %>%
  gather(CellType, EigenGeneExpression, -ID, -Diagnosis, -Tissue) %>%
  ggplot(aes(x = Diagnosis, y = abs(EigenGeneExpression), fill = Tissue)) + geom_boxplot() + facet_grid(.~CellType)
p = p +  theme(axis.text.x = element_text(angle = 90, hjust = 1))   
p
```

@@

pl = list()
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=factor(Source), color=as.factor(Tissue), size=as.numeric(OPC)))
p <- p + theme_bw() + theme(legend.position="top") +scale_shape_manual(values = 1:2)
# p <- p + geom_text(aes(label= ID), size=4, hjust=0)
pl[[1]] <- p + xlab(paste0('PC1(', round(PC$sdev[1]^2/sum(PC$sdev^2)*100), '%)')) + ylab(paste0('PC2(', round(PC$sdev[2]^2/sum(PC$sdev^2)*100), '%)'))

p <- ggplot(plotdata, aes(x=PC3, y=PC4))
p <- p + geom_point(aes(shape=factor(Source), color=as.factor(Diagnosis), size=as.numeric(OPC)))
p <- p + theme_bw() + theme(legend.position="top") +scale_shape_manual(values = 1:2)
# p <- p + geom_text(aes(label= ID), size=4, hjust=0)
pl[[2]] <- p + xlab(paste0('PC3(', round(PC$sdev[3]^2/sum(PC$sdev^2)*100), '%)')) + ylab(paste0('PC4(', round(PC$sdev[4]^2/sum(PC$sdev^2)*100), '%)'))

multiplot(plotlist = pl, cols = 2)
```
```{r decompse.normalise.data2, fig.height=9, fig.width=25, results='asis'}
# Tree based clustering
sampleTrees = flashClust::hclust(dist(t(VOOM.GENE_EXPRESSION$E)), method = "average")
tree = WGCNA::cutreeStatic(sampleTrees, cutHeight = 135, minSize=10)

@@

### Clustering residual gene expression
```{r decompse.normalise.data3, fig.height=15, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION[,-(1)], scale.=T)

@@

# Get differnetial expression
DE = list()
p = list()
for (i in 1:12){
  DE[[i]] = topTable(FIT.CONTR, coef=i, number = dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1])
  DE[[i]] = rownameToFirstColumn(DE[[i]], 'ensembl_gene_id') %>%
    left_join(Ensemble2HGNC)
  p[[i]] <- ggplot(DE[[i]], aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
  p[[i]] <- p[[i]] + geom_hline(yintercept = -log10(0.05), color = 'red') 
  p[[i]] <- p[[i]] + geom_vline(xintercept = -log2(1.5), color = 'red')
  p[[i]] <- p[[i]] + geom_vline(xintercept = log2(1.5), color = 'red')
  p[[i]] <- p[[i]] + ggtitle(gsub('BrainRegion.Diagnosis','',colnames(contrast)[i]))
}
multiplot(plotlist=p, cols = 3)
names(DE) = gsub('BrainRegion.Diagnosis','',colnames(contrast))

DE = plyr::ldply(DE, .id = 'Comparison')
tmp = DE %>% 
  dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.5)) %>%
  group_by(Comparison) %>%
  dplyr::summarize(nDiffExpGenes = length(unique(hgnc_symbol)))
kable(tmp)
```
